\vspace{-2cm} Inference on the Effects of Observed Features \\ in Latent Variable Models for Networks

### 2018-04-23 manuscript_rearranged.tex
clarified explanation of covariate generation

%============Article Title, Authors==================
\title{\vspace{-2cm} Inference on the Effects of Observed Features
\\ in Latent Space Models for Networks }

@@

For simplicity and computational efficiency we consider a one-dimensional latent space. That is, we create a variable that can be represented using the Euclidean distance between two nodes in a one-dimensional space. We expect our results to generalize to higher dimensions and other variants of the LSM since the tendency for latent variables to be used in explaining structure that cannot be attributed covariates is universal with respect to both the dimension of the latent space and the form of the LSM. To generate an observed covariate that exhibits a controllable degree of correlation with the latent network structure we follow a four-step process. First, we simulate one-dimensional positions for each node, drawing from a standard normal distribution, and calculate the Euclidean distance (i.e., absolute distance) $\mathbf{d}$ between each pair of positions. Second, we generate a target covariance matrix ($\Sigma$) using the $c$-vine method of \cite{lewandowski2009generating} [UNPACK FROM https://www.rdocumentation.org/packages/clusterGeneration/versions/1.3.4/topics/genPositiveDefMat] which generates a uniform draw from the space of positive-definite matrices. The $c$-vine method makes use of a parameter $\eta$, which when equal to $1$ generates draws from uniformly from the space of positive-definite matrices, making the correlation between the latent positions and the measured covariate 0 in expectation. As $\eta$ increases above $1$ draws of $\Sigma$ have increasingly weak correlation\footnote{We show the distribution of correlations under each value of $\eta$ used in our simulation in Figure \ref{fig:vine}.}. Third, given the target covariance matrix ($\Sigma$) among the covariates and distances, $\langle \mathbf{x}, \mathbf{d} \rangle$, we derive the conditional mean vector and covariance, assuming that $\mathbf{x}$ has a normal distribution given $\mathbf{d}$ (see \cite[pp. 116--117]{eaton1983} for the conditional normal derivation).
+Fourth, we simulate $x$ as a normal random variable with the respective conditional means and covariance. Finally we standardize $x$ to have zero mean and unit variance to ensure that the scale of estimation errors is comparable across simulations.

@@

\begin{figure}
\includegraphics[width=\textwidth]{figures/max_r_vine.png}
\caption{The distribution of the absolute value of the correlation between the observed covariate and the latent distances. \label{fig:vine}}
\end{figure}

>>>>

%============Article Title, Authors==================
\title{\vspace{-2cm} Inference on the Effects of Observed Features
\\ in Latent Variable Models for Networks }

@@

For simplicity and computational efficiency we consider a one-dimensional latent space. That is, we create a variable that can be represented using the Euclidean distance between two nodes in a one-dimensional space. We expect our results to generalize to higher dimensions and other variants of the LSM since the tendency for latent variables to be used in explaining structure that cannot be attributed covariates is universal with respect to both the dimension of the latent space and the form of the LSM. To generate an observed covariate that exhibits a controllable degree of correlation with the latent network structure we follow a four-step process. The steps are as follows.
\begin{enumerate}
\item  Simulate one-dimensional positions for each node (i.e., a scalar value for each node), drawing from a standard normal distribution, and calculate the Euclidean distance (i.e., absolute distance) $\mathbf{d}$ between each pair of positions. We then normalize $\mathbf{d}$ to assure it has unit variance by $\mathbf{d} := \mathbf{d}/ \text{sd}(\mathbf{d}),$ where sd($\cdot$) is the empirical standard deviation.

 \item Generate a correlation coefficient $(\rho)$ on the interval $(-1,1)$ as $\rho = 1-2\times u$, where $u \sim  \text{Beta}(\eta,\eta)$. $\eta$ is used to control the magnitude of the correlation coefficient. The smaller the value of $\eta$, the larger the absolute magnitude of the correlation coefficient, in expectation. This relationship is illustrated in Figure \ref{fig:vine}. When $\eta = 1$, the correlations generated exhibit the expected uniform distribution. When $\eta = 1,000,000$, each correlation coefficient is approximately 0. \footnote{This methodology generalizes to higher dimensional correlation matrices as the `c-vine' method developed by \cite{lewandowski2009generating}, and implemented in the \R package \texttt{clusterGeneration} \citep{clusterGeneration}, which we use in our simulations.}

 \item Simulate $\mathbf{x} = \frac{\rho}{\sqrt{1-\rho^2}}\mathbf{d}+\mathbf{w},$ where $\mathbf{w} \sim N(0,1)$. This assures that the correlation between $\mathbf{x}$ and $\mathbf{d}$ is $\rho$.\footnote{This could be generalized to more than one covariate using the conditional normal distribution derived by \cite[pp. 116--117]{eaton1983}.}

 \item Finally we standardize $\mathbf{x} := \mathbf{x}/ \text{sd}(\mathbf{x})$ to have unit variance to ensure that the variance of $\mathbf{x}$ relative to $\mathbf{d}$ is unrelated to $\rho$. This is to assure that $\rho$ does not affect the efficiency with which the effect of $\mathbf{x}$ can be estimated through some property unrelated to the relationship between $\mathbf{x}$ and $\mathbf{d}$.

\end{enumerate}

@@

\begin{figure}
\includegraphics[width=\textwidth]{figures/max_r_vine.png}
\caption{The distribution of the absolute value of the correlation between the observed covariate and the latent distances. The value in the plot title corresponds to the value of $\eta$. \label{fig:vine}}
\end{figure}
